---
title: "Age and sex-related variability in the presentation of anxiety and depression symptoms in the GLAD Study - highest education post hoc analysis"
author: "Katherine N Thompson"
date: "16/06/2021"
output:
  word_document:
    toc: yes
    toc_depth: '2'
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: no
    number_sections: no
    highlight: monochrome
    theme: cerulean
code_folding: show
html_notebook:
  theme: cerulean
toc: yes
---


# Set up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      comment = NA,
                      prompt = FALSE,
                      cache = FALSE)
```

Delete everything in your global environment
```{r Clear global environment, include=FALSE}
remove(list = ls())
```

Packages
```{r Load packages, include=FALSE}
library(knitr)
library(car)
library(psych)
library(lubridate)
library(broom)
library(summarytools)
library(data.table)
library(eeptools)
library(magrittr)
library(reshape2)
library(lavaan)
library(mlogit)
library(foreign)
library(sjPlot)
library(polycor)
library(REdaS)
library(scales)
library(forcats)
library(GPArotation)
library(tidyverse)
```

Retrieve the recent date
```{r Recent date, include=FALSE}
date = Sys.Date()
```

Define colours for plotting
```{r GLAD colour palette , include=FALSE}
palette = c(
  "#efc00b", #female
  "#b7dee8"  #male
  )

palette3 <- c("#78D9C5","#F5BE5E","#EEB6E9")
```

```{r source data path file}
source("../data_paths.R")
```

## Prepare item lists
```{r prepare item groups, include=FALSE}
#All the items in numeric form
ALL.items.full_numeric <- c("phq.anhedonia_numeric",
             "phq.depressed_mood_numeric",
             "phq.sleep_change_numeric",
             "phq.little_energy_numeric",
             "phq.eating_problems_numeric",
             "phq.worthlessness_numeric",
             "phq.concentration_problems_numeric",
             "phq.motor_problems_numeric",
             "phq.suicide_ideation_numeric",
             "gad.feeling_anxious_numeric",
             "gad.control_worrying_numeric",
             "gad.worry_toomuch_numeric",
             "gad.trouble_relaxing_numeric", 
             "gad.restless_numeric", 
             "gad.irritable_numeric",
             "gad.feeling_afraid_numeric")

#All the items as factors
ALL.items.full <- c("phq.anhedonia",
             "phq.depressed_mood",
             "phq.sleep_change",
             "phq.little_energy",
             "phq.eating_problems",
             "phq.worthlessness",
             "phq.concentration_problems",
             "phq.motor_problems",
             "phq.suicide_ideation",
             "gad.feeling_anxious",
             "gad.control_worrying",
             "gad.worry_toomuch",
             "gad.trouble_relaxing", 
             "gad.restless", 
             "gad.irritable",
             "gad.feeling_afraid")

# Create vector with items that are included in total score
GAD.items = c("gad.feeling_anxious_numeric",
              "gad.control_worrying_numeric",
              "gad.worry_toomuch_numeric",
              "gad.trouble_relaxing_numeric",
              "gad.restless_numeric",
              "gad.irritable_numeric",
              "gad.feeling_afraid_numeric")

# Create vector with items that are included in total score
PHQ.items = c("phq.anhedonia_numeric",
              "phq.depressed_mood_numeric",
              "phq.sleep_change_numeric",
              "phq.little_energy_numeric",
              "phq.eating_problems_numeric",
              "phq.worthlessness_numeric",
              "phq.concentration_problems_numeric",
              "phq.motor_problems_numeric",
              "phq.suicide_ideation_numeric")

#list of categorical items
cat.items <- c("phq.anhedonia.cat",
             "phq.depressed_mood.cat",       
             "phq.sleep_change.cat",
             "phq.little_energy.cat",
             "phq.eating_problems.cat",
             "phq.worthlessness.cat",
             "phq.concentration_problems.cat",
             "phq.motor_problems.cat",
             "phq.suicide_ideation.cat",
             "gad.feeling_anxious.cat",
             "gad.control_worrying.cat",
             "gad.worry_toomuch.cat",
             "gad.trouble_relaxing.cat", 
             "gad.restless.cat", 
             "gad.irritable.cat",
             "gad.feeling_afraid.cat")

cat.items.phq <- c("phq.anhedonia.cat",
             "phq.depressed_mood.cat",       
             "phq.sleep_change.cat",
             "phq.little_energy.cat",
             "phq.eating_problems.cat",
             "phq.worthlessness.cat",
             "phq.concentration_problems.cat",
             "phq.motor_problems.cat",
             "phq.suicide_ideation.cat")

cat.items.gad <- c(
             "gad.feeling_anxious.cat",
             "gad.control_worrying.cat",
             "gad.worry_toomuch.cat",
             "gad.trouble_relaxing.cat", 
             "gad.restless.cat", 
             "gad.irritable.cat",
             "gad.feeling_afraid.cat")



#categorical items labeled for endorsement plot 
cat.items.labelled <- c("phq.anhedonia.cat.labelled",
             "phq.depressed_mood.cat.labelled",       
             "phq.sleep_change.cat.labelled",
             "phq.little_energy.cat.labelled",
             "phq.eating_problems.cat.labelled",
             "phq.worthlessness.cat.labelled",
             "phq.concentration_problems.cat.labelled",
             "phq.motor_problems.cat.labelled",
             "phq.suicide_ideation.cat.labelled",
             "gad.feeling_anxious.cat.labelled",
             "gad.control_worrying.cat.labelled",
             "gad.worry_toomuch.cat.labelled",
             "gad.trouble_relaxing.cat.labelled", 
             "gad.restless.cat.labelled", 
             "gad.irritable.cat.labelled",
             "gad.feeling_afraid.cat.labelled")

#create data frame that doesn't include "worrying too much" item
ALL.items.full.withoutworry <- c("phq.anhedonia",
                                 "phq.depressed_mood",
                                 "phq.sleep_change",
                                 "phq.little_energy",
                                 "phq.eating_problems",
                                 "phq.worthlessness",
                                 "phq.concentration_problems",
                                 "phq.motor_problems",
                                 "phq.suicide_ideation",
                                 "gad.feeling_anxious",
                                 "gad.control_worrying",
                                 "gad.trouble_relaxing", 
                                 "gad.restless", 
                                 "gad.irritable",
                                 "gad.feeling_afraid")

#create data frame that doesn't include "concentration problems" item
ALL.items.full.withoutconcentration <- c("phq.anhedonia",
                                 "phq.depressed_mood",
                                 "phq.sleep_change",
                                 "phq.little_energy",
                                 "phq.eating_problems",
                                 "phq.worthlessness",
                                 "phq.motor_problems",
                                 "phq.suicide_ideation",
                                 "gad.feeling_anxious",
                                 "gad.control_worrying",
                                 "gad.trouble_relaxing", 
                                 "gad.restless", 
                                 "gad.irritable",
                                 "gad.feeling_afraid")

#create data frame that doesn't include "concentration problems" and "irritability" item
ALL.items.full.withoutirritability <- c("phq.anhedonia",
                                 "phq.depressed_mood",
                                 "phq.sleep_change",
                                 "phq.little_energy",
                                 "phq.eating_problems",
                                 "phq.worthlessness",
                                 "phq.motor_problems",
                                 "phq.suicide_ideation",
                                 "gad.feeling_anxious",
                                 "gad.control_worrying",
                                 "gad.trouble_relaxing", 
                                 "gad.restless",
                                 "gad.feeling_afraid")

#continuous variables for descriptives
continuous.variables <- c("PHQ.total.score", 
                          "GAD.total.score", 
                          "age")
```

```{r read in data and check variable names, include=FALSE}
dat <- readRDS(file = paste0(data_path, "data_formatted_FA.rds"))
colnames(dat)
```

## Create data sets with just men/women
```{r datasets for males and males}
data.female <- dat %>% filter(sex == "Female")
data.male <- dat %>% filter (sex == "Male")
```

# **Age and sex-related variability in the presentation of PHQ and GAD items**

### Create age/10 variable 
```{r divide age by 10}
#this makes the OR more interpretable as you think of increase/decrease per 10 years rather than per year
dat <- dat %>%
  mutate(
    age_divided = 
      (age)/10
  )
colnames(dat)
```

## Logistic regressions for age
### Logistic regressions for PHQ items - age divided, sex, PHQ total score, GAD total score 

```{r create loop to run all logistic regressions for phq items}
logistic.phq.stats.sex <- list() # create empty list to put statistics into
logistic.phq.stats.age <- list() # create empty list to put statistics into

for (i in 1:length(cat.items.phq))
{ cat(i)
  # create the logistic regression formula to be used for each categorical phq9 item
  logistic.formula.phq <- as.formula(sprintf("%s ~ age_divided + sex + PHQ.total.score + GAD.total.score + highest_education", cat.items.phq[i]))
  
  # run the regression
  logistic.reg.phq <- glm(logistic.formula.phq, data = dat, family = "binomial")
  
  # call the tidy output for sex
  logistic.reg.phq.output.sex <- broom::tidy(logistic.reg.phq, conf.int = TRUE, exponentiate = TRUE, conf.level = 0.95)[-c(1,2,4:8),] %>% #remove all rows except sex
    dplyr::rename(
           OR = estimate,
           OR.CI.lower = conf.low,
           OR.CI.upper = conf.high)
  
  # call the tidy output for age
  logistic.reg.phq.output.age <- broom::tidy(logistic.reg.phq, conf.int = TRUE, exponentiate = TRUE, conf.level = 0.95)[-c(1,3:8),] %>% #remove all rows except age
    dplyr::rename(
           OR = estimate,
           OR.CI.lower = conf.low,
           OR.CI.upper = conf.high)
  
  # feed it back into the original empty lists
  logistic.phq.stats.sex[[i]] <- logistic.reg.phq.output.sex
  logistic.phq.stats.age[[i]] <- logistic.reg.phq.output.age
  
}

# set the names of the list to match the variable names
names(logistic.phq.stats.sex) <- cat.items.phq
names(logistic.phq.stats.age) <- cat.items.phq
```

## Call statistics from loop

```{r call statistics for all phq regressions, include=FALSE}
#sex
for (i in 1:length(cat.items.phq))
{ 
  assign(paste0("log.", names(logistic.phq.stats.sex[i]), ".sex"), as.tibble(logistic.phq.stats.sex[[i]]))
}

#age
for (i in 1:length(cat.items.phq))
{ 
  assign(paste0("log.", names(logistic.phq.stats.age[i]), ".age"), as.tibble(logistic.phq.stats.age[[i]]))
}
```

Join table rows together from the PHQ regression stats 
```{r join datasets to make forest plot}
# bind all the PHQ/age rows together
ALL_CI_PHQ.age <- bind_rows(log.phq.anhedonia.cat.age, 
                            log.phq.depressed_mood.cat.age, 
                            log.phq.sleep_change.cat.age, 
                            log.phq.little_energy.cat.age, 
                            log.phq.eating_problems.cat.age, 
                            log.phq.worthlessness.cat.age, 
                            log.phq.concentration_problems.cat.age, 
                            log.phq.motor_problems.cat.age, 
                            log.phq.suicide_ideation.cat.age)

# add a column to show the disorder and symptom names
ALL_CI_PHQ.age <- ALL_CI_PHQ.age %>%
  mutate(disorder = "MDD") %>%
  mutate(symptom = c("Anhedonia", 
                     "Depressed mood", 
                     "Sleep change", 
                     "Little energy", 
                     "Weight or appetite problems", 
                     "Worthlessness", 
                     "Concentration problems", 
                     "Motor problems",
                     "Suicide ideation"))

#bind all the PHQ/sex rows together
ALL_CI_PHQ.sex <- bind_rows(log.phq.anhedonia.cat.sex, 
                            log.phq.depressed_mood.cat.sex, 
                            log.phq.sleep_change.cat.sex, 
                            log.phq.little_energy.cat.sex, 
                            log.phq.eating_problems.cat.sex, 
                            log.phq.worthlessness.cat.sex, 
                            log.phq.concentration_problems.cat.sex, 
                            log.phq.motor_problems.cat.sex, 
                            log.phq.suicide_ideation.cat.sex)

#add a column to show the disorder
ALL_CI_PHQ.sex <- ALL_CI_PHQ.sex %>%
  mutate(disorder = "MDD") %>%
  mutate(symptom = c("Anhedonia", 
                     "Depressed mood", 
                     "Sleep change", 
                     "Little energy", 
                     "Weight or appetite problems", 
                     "Worthlessness", 
                     "Concentration problems", 
                     "Motor problems",
                     "Suicide ideation"))
```

```{r create loop to run all logistic regressions for gad items}
logistic.gad.stats.sex <- list() # create empty list to put statistics into
logistic.gad.stats.age <- list() # create empty list to put statistics into

for (i in 1:length(cat.items.gad))
{ cat(i)
  # create the logistic regression formula to be used for each categorical gad9 item
  logistic.formula.gad <- as.formula(sprintf("%s ~ age_divided + sex + PHQ.total.score + GAD.total.score + highest_education", cat.items.gad[i]))
  
  # run the regression
  logistic.reg.gad <- glm(logistic.formula.gad, data = dat, family = "binomial")
  
  # call the tidy output for sex
  logistic.reg.gad.output.sex <- broom::tidy(logistic.reg.gad, conf.int = TRUE, exponentiate = TRUE, conf.level = 0.95)[-c(1,2,4:8),] %>% #remove all rows except sex
    dplyr::rename(
           OR = estimate,
           OR.CI.lower = conf.low,
           OR.CI.upper = conf.high)
  
  # call the tidy output for age
  logistic.reg.gad.output.age <- broom::tidy(logistic.reg.gad, conf.int = TRUE, exponentiate = TRUE, conf.level = 0.95)[-c(1,3:8),] %>% #remove all rows except age
    dplyr::rename(
           OR = estimate,
           OR.CI.lower = conf.low,
           OR.CI.upper = conf.high)
  
  # feed it back into the original empty lists
  logistic.gad.stats.sex[[i]] <- logistic.reg.gad.output.sex
  logistic.gad.stats.age[[i]] <- logistic.reg.gad.output.age
  
}

# set the names of the list to match the variable names
names(logistic.gad.stats.sex) <- cat.items.gad
names(logistic.gad.stats.age) <- cat.items.gad
```

## Call statistics from loop

```{r call statistics for all gad regressions, include=FALSE}
#sex
for (i in 1:length(cat.items.gad))
{ 
  assign(paste0("log.", names(logistic.gad.stats.sex[i]), ".sex"), as.tibble(logistic.gad.stats.sex[[i]]))
}

#age
for (i in 1:length(cat.items.gad))
{ 
  assign(paste0("log.", names(logistic.gad.stats.age[i]), ".age"), as.tibble(logistic.gad.stats.age[[i]]))
}
```

Combine the data together
```{r combine data for all gad symptoms}
#bind all the GAD/age rows together
ALL_CI_GAD.age <- bind_rows(log.gad.feeling_anxious.cat.age, 
                            log.gad.control_worrying.cat.age, 
                            log.gad.worry_toomuch.cat.age, 
                            log.gad.trouble_relaxing.cat.age, 
                            log.gad.restless.cat.age, 
                            log.gad.irritable.cat.age, 
                            log.gad.feeling_afraid.cat.age)
#add in column to show the disorder
ALL_CI_GAD.age <- ALL_CI_GAD.age %>%
  mutate(disorder = "GAD") %>%
  mutate(symptom = c("Feeling anxious", 
                     "Difficulty controlling worrying", 
                     "Worrying too much", 
                     "Trouble relaxing", 
                     "Restlessness", 
                     "Irritable", 
                     "Feeling afraid"))

#bind all the GAD/sex rows together
ALL_CI_GAD.sex <- bind_rows(log.gad.feeling_anxious.cat.sex, 
                            log.gad.control_worrying.cat.sex, 
                            log.gad.worry_toomuch.cat.sex, 
                            log.gad.trouble_relaxing.cat.sex, 
                            log.gad.restless.cat.sex, 
                            log.gad.irritable.cat.sex, 
                            log.gad.feeling_afraid.cat.sex)
#add in column to show the disorder
ALL_CI_GAD.sex <- ALL_CI_GAD.sex %>%
  mutate(disorder = "GAD") %>%
  mutate(symptom = c("Feeling anxious", 
                     "Difficulty controlling worrying", 
                     "Worrying too much", 
                     "Trouble relaxing", 
                     "Restlessness", 
                     "Irritable", 
                     "Feeling afraid"))
```


Bind all the rows of the tables together for age and sex, and then order according to OR
```{r combine all symptoms and reorder them}
#bind all age rows together
ALL_CI.age <- bind_rows(ALL_CI_GAD.age, ALL_CI_PHQ.age)

ALL_CI.age <- ALL_CI.age %>%
  mutate(p.value.adjusted = 
           p.adjust(p.value, 
                    method = "fdr",
                    n = 16)) %>% #calculated separately at only doing 16 comparisons
  mutate(term = "Age/10") # rename the term
 
#order
ALL_CI.age.ordered <- arrange(ALL_CI.age, desc(OR))

#bind all sex rows together
ALL_CI.sex <- bind_rows(ALL_CI_GAD.sex, ALL_CI_PHQ.sex)

ALL_CI.sex <- ALL_CI.sex %>%
  mutate(p.value.adjusted = 
           p.adjust(p.value, 
                    method = "fdr",
                    n = 16))  %>% #calculated separately at only doing 16 comparisons
 mutate(term = "Sex (Female)") 

#order
ALL_CI.sex.ordered <- arrange(ALL_CI.sex, desc(OR))

#bind all rows together 
ALL_CI <- bind_rows(ALL_CI.age, ALL_CI.sex)

#add significance variable
ALL_CI <- ALL_CI %>%
  mutate(
    Significant.raw = 
      if_else(
        p.value.adjusted < 0.05, 
        1,
        0
      )
  )
#turn into factor
ALL_CI <- ALL_CI %>%
  mutate(Significant = 
  recode_factor(Significant.raw,
    `1` = "Significant",
    `0` = "Non-significant"
      )
  )

#order
ALL_CI.ordered <- arrange(ALL_CI, desc(OR)) #ordered by age variables 
```

Make the symptom into a factor with levels that reflect the PHQ and GAD symptoms
```{r reorder levels of symptoms}
levels <- paste(ALL_CI.sex.ordered$symptom)
#make symptom variable in age data table
ALL_CI.age.ordered$symptom <- factor(ALL_CI.age.ordered$symptom,
                                     levels = paste(ALL_CI.age.ordered$symptom,
                                                    sep = ","))
#make symptom variable in sex data table
ALL_CI.sex.ordered$symptom <- factor(ALL_CI.sex.ordered$symptom,
                                     levels = paste(ALL_CI.sex.ordered$symptom,
                                                    sep = ","))

#make symptom variable in sex data table
ALL_CI.ordered$symptom <- factor(ALL_CI.ordered$symptom,
                                     levels = paste(levels,
                                                    sep = ","))

kable(ALL_CI.ordered) 
```

## Age and sex plots {.tabset .tabset-fade}

## Combined age and sex related variability 
```{r joint ggplot of age and sex variation}
combined_symptoms_plot <- ggplot(
  ALL_CI.ordered,
  aes(x = symptom,
      y = OR,
      ymin = OR.CI.lower,
      ymax = OR.CI.upper,
      colour = term)
  ) +
  scale_fill_manual(values = palette3) +
  scale_color_manual(values = palette3) +
  scale_shape_manual(values = c(19, 21)) +
  geom_pointrange(aes(shape = Significant), position = position_dodge(width = 0.7)) +
  labs(y = "OR (95% CI)",
       x = "Symptom",
       title = "Age and sex-related variability in PHQ-9 and GAD-7 symptoms",
       subtitle = paste("N(Total) = ", length(dat$PHQ.total.score),
                        "; N(Female) = ", sum(!is.na(dat$PHQ.total.score[dat$sex=="Female"])),
                        "; N(Male) = ", sum(!is.na(dat$PHQ.total.score[dat$sex=="Male"])), sep = ""),
       color = "",
       shape = "") +
  theme(panel.grid.major.x = element_line(size = 0.5,
                                          linetype = 'dashed',
                                          colour = "gray"),
        plot.title = element_text(size = 15, face = "bold", hjust=0.5),
        axis.title.x = element_text(size=12,face="bold"),
        axis.title.y = element_blank(),
        axis.text.x = element_text(colour = "black", size = 9),
        axis.text.y = element_text(colour = "black", size = 12),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        panel.background = element_blank(),
        legend.position = "bottom") +
  geom_hline(linetype = "dashed", yintercept = 1) +
  scale_y_continuous(trans = log10_trans(), breaks=c(0.4, 0.6, 0.8, 1, 1.2, 1.4, 1.6, 1.8, 2)) +
  coord_flip()

combined_symptoms_plot
```

# **Age and sex related variability for the identified factors**

## Create correlation matrix of each sum score and factor score 

```{r create data frame for factors correlation matrix}
factor.sum.score.items <- c("PHQ.total.score",
                           "GAD.total.score",
                           "worry",
                           "mood",
                           "somatic",
                           "motor")

factor.sum.score.df <- as.data.frame(dat [,factor.sum.score.items])  #create your dataframe with the sum scores and factor scores 

is.data.frame(factor.sum.score.df)  #check that it is a dataframe 

colnames(factor.sum.score.df) <- c("PHQ sum score",   #Add in the names you would like for your variables
                                 "GAD sum score", 
                                 "Worry factor score", 
                                 "Mood factor score", 
                                 "Somatic factor score",
                                 "Motor factor score")
```

```{r correlation matrix for factor scores and sum scores}
factor.sum.score.matrix <- cor(factor.sum.score.df, 
                               method = "pearson", 
                               use = "complete.obs")
factor.sum.score.matrix 
```

```{r heat map for factor and sum scores}
# Reorder the correlation matrix
cor_matrix_factorsum <- reorder_cor_matrix(factor.sum.score.matrix)
  
#melt the values
metled_cor_matrix_factorsum <- melt(cor_matrix_factorsum, na.rm = TRUE)
metled_cor_matrix_factorsum

#correlation heat map
factorsum_correlation_heat_map <- ggplot(metled_cor_matrix_factorsum, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white",
   midpoint = 0, limit = c(-1,1), space = "Lab",
    name="Pearson\nCorrelation") +
  theme_minimal() +
  labs(y = "Items",
       x = "Items",
       title = "Correlation heat map for factor scores and sum scores"
       )+
 theme(axis.text.x = element_text(angle = 45, vjust = 1,
    size = 12, hjust = 1))+
 coord_fixed() +
  geom_text(aes(Var2, Var1, label = round(value, digits = 2)), color = "black", size = 3)

factorsum_correlation_heat_map
```

## Linear regressions for factors - age_divided + sex + PHQ.total.score + GAD.total.score + highest_education

R-squared is a goodness-of-fit measure for linear regression models. 
This statistic indicates the percentage of the variance in the dependent variable that the independent variables explain collectively. 
```{r linear regression models for factor scores}
#Worry factor scores
lin.worry_factor <- lm(worry ~ age_divided + sex + PHQ.total.score + GAD.total.score + highest_education,
              data = dat)
summary(lin.worry_factor ) #standard summary of model
CI.lin.worry_factor <- tidy(lin.worry_factor, conf.int = TRUE, exponentiate = TRUE) #tidy summary with confidence intervals
CI.lin.worry_factor <- add_column(CI.lin.worry_factor, #add column to indicate the factor
  factor = "Worry symptoms")
CI.lin.worry_factor #previouslt used OR.log.motor_factor <- exp(coef(log.motor_factor.endorsement)) to get OR
CI.lin.worry_factor.age <- CI.lin.worry_factor[2,1:8]  
CI.lin.worry_factor.age
CI.lin.worry_factor.sex <- CI.lin.worry_factor[3,1:8] 
CI.lin.worry_factor.sex

#Mood factor scores
lin.mood_factor <- lm(mood ~ age_divided + sex + PHQ.total.score + GAD.total.score + highest_education,
              data = dat)
summary(lin.mood_factor ) #standard summary of model
CI.lin.mood_factor <- tidy(lin.mood_factor, conf.int = TRUE, exponentiate = TRUE) #tidy summary with confidence intervals
CI.lin.mood_factor <- add_column(CI.lin.mood_factor, #add column to indicate the factor
  factor = "Mood symptoms")
CI.lin.mood_factor 
CI.lin.mood_factor.age <- CI.lin.mood_factor[2,1:8]  
CI.lin.mood_factor.age
CI.lin.mood_factor.sex <- CI.lin.mood_factor[3,1:8]  
CI.lin.mood_factor.sex

#Somatic factor scores
lin.somatic_factor <- lm(somatic ~ age_divided + sex + PHQ.total.score + GAD.total.score + highest_education,
              data = dat)
summary(lin.somatic_factor ) #standard summary of model
CI.lin.somatic_factor <- tidy(lin.somatic_factor, conf.int = TRUE, exponentiate = TRUE) #tidy summary with confidence intervals
CI.lin.somatic_factor <- add_column(CI.lin.somatic_factor, #add column to indicate the factor
  factor = "Somatic symptoms")
CI.lin.somatic_factor 
CI.lin.somatic_factor.age <- CI.lin.somatic_factor[2,1:8]  
CI.lin.somatic_factor.age
CI.lin.somatic_factor.sex <- CI.lin.somatic_factor[3,1:8]  
CI.lin.somatic_factor.sex

#Motor factor scores
lin.motor_factor <- lm(motor ~ age_divided + sex + PHQ.total.score + GAD.total.score + highest_education,
              data = dat)
summary(lin.motor_factor ) #standard summary of model
CI.lin.motor_factor <- tidy(lin.motor_factor, conf.int = TRUE, exponentiate = TRUE) #tidy summary with confidence intervals
CI.lin.motor_factor <- add_column(CI.lin.motor_factor, #add column to indicate the factor
  factor = "Motor symptoms")
CI.lin.motor_factor 
CI.lin.motor_factor.age <- CI.lin.motor_factor[2,1:8]  
CI.lin.motor_factor.age
CI.lin.motor_factor.sex <- CI.lin.motor_factor[3,1:8]  
CI.lin.motor_factor.sex
```

Combine rows of CIs
```{r combine datasets }
#bind all age rows together
ALL_CI_factor.age <- bind_rows(CI.lin.mood_factor.age, 
                               CI.lin.worry_factor.age, 
                               CI.lin.somatic_factor.age, 
                               CI.lin.motor_factor.age)
#pvalue correction
ALL_CI_factor.age <- ALL_CI_factor.age %>%
  mutate(p.value.adjusted = 
           p.adjust(p.value, 
                    method = "fdr",
                    n = 8)) #calculated separately at only doing 8comparisons
#rename term
ALL_CI_factor.age  <- ALL_CI_factor.age %>%
  mutate(
    term = 
      "Age/10"
  )

#order
ALL_CI_factor.age.ordered <- 
  arrange(ALL_CI_factor.age, desc(estimate))


#bind all sex rows together
ALL_CI_factor.sex <- bind_rows(CI.lin.mood_factor.sex, 
                               CI.lin.worry_factor.sex, 
                               CI.lin.somatic_factor.sex, 
                               CI.lin.motor_factor.sex)
#adjust pvalue
ALL_CI_factor.sex <- ALL_CI_factor.sex %>%
  mutate(p.value.adjusted = 
           p.adjust(p.value, 
                    method = "fdr",
                    n = 16)) #calculated separately at only doing 16 comparisons
#rename term
ALL_CI_factor.sex <- ALL_CI_factor.sex %>%
  mutate(
    term = 
      "Sex (Female)"
  )
#order
ALL_CI_factor.sex.ordered <- 
  arrange(ALL_CI_factor.sex, desc(estimate))

#bind all rows together 
ALL_CI_factor <- bind_rows(ALL_CI_factor.age, ALL_CI_factor.sex)
ALL_CI_factor 

#add significance variable
ALL_CI_factor <- ALL_CI_factor %>%
  mutate(
    Significant.raw = 
      if_else(
        p.value.adjusted < 0.05, 
        1,
        0
      )
  )
#turn into factor
ALL_CI_factor <- ALL_CI_factor %>%
  mutate(Significant = 
  recode_factor(Significant.raw,
    `1` = "Significant",
    `0` = "Non-significant"
      )
  )

#order
ALL_CI_factor.ordered <- arrange(ALL_CI_factor, desc(estimate)) #ordered by age variables 
kable(ALL_CI_factor.ordered)
```

Make the symptom factor variable into a factor
```{r make the factor variable into a factor}
levels_factor <- ALL_CI_factor.sex.ordered$factor

ALL_CI_factor.age.ordered$factor <- factor(ALL_CI_factor.age.ordered$factor, 
                                           levels = paste(ALL_CI_factor.age.ordered$factor, 
                                                          sep = ","))

ALL_CI_factor.sex.ordered$factor <- factor(ALL_CI_factor.sex.ordered$factor, 
                                           levels = paste(ALL_CI_factor.sex.ordered$factor, 
                                                          sep = ","))

ALL_CI_factor.ordered$factor <- factor(ALL_CI_factor.ordered$factor, 
                                           levels = paste(levels_factor, 
                                                          sep = ","))

```

## Factor plots  {.tabset .tabset-fade}

### Combined 
```{r combined factors plot}
combined_factors_plot <- ggplot(
  ALL_CI_factor.ordered,
  aes(x = factor,
      y = estimate,
      ymin = conf.low,
      ymax = conf.high,
      colour = term)
  ) +
  scale_fill_manual(values = palette3) +
  scale_color_manual(values = palette3) + 
  scale_shape_manual(values = c(19, 21)) +
  geom_pointrange(aes(shape = Significant), position = position_dodge2(width = 0.3), size = 0.8) +
  labs(y = "Standardised Estimate (95% CI)",
       x = "Symptom",
       title = "Age and sex-related variability in anxiety and depression factors",
       subtitle = paste("N(Total) = ", length(dat$PHQ.total.score),
                        "; N(Female) = ", sum(!is.na(dat$PHQ.total.score[dat$sex=="Female"])),
                        "; N(Male) = ", sum(!is.na(dat$PHQ.total.score[dat$sex=="Male"])), sep = ""),
       color = "",
       shape = "") +
  theme(panel.grid.major.x = element_line(size = 0.5,
                                          linetype = 'dashed',
                                          colour = "gray"),
        plot.title = element_text(size = 15, face = "bold", hjust = 0.5),
        axis.title.x = element_text(size = 12, face = "bold"),
        axis.title.y = element_blank(),
        axis.text.x = element_text(colour = "black", size = 9),
        axis.text.y = element_text(colour = "black", size = 12),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        panel.background = element_blank(),
        legend.position = "bottom") +
  geom_hline(linetype = "dashed", yintercept = 0) +
 coord_flip()
combined_factors_plot
```

# Panel plots

## Age and sex variation in symptoms and factors
```{r panel plots for factor and symptoms}
library(ggpubr)
combined_symptoms_plot <- combined_symptoms_plot + 
  labs(title = "Symptom level") + 
  theme(plot.title = element_text(size = 12,face = "bold"),
        axis.title.x = element_text(size = 10, face = "bold"))

combined_factors_plot <- combined_factors_plot + 
  labs(title = "Factor level") + 
  theme(plot.title = element_text(size = 12,face = "bold"),
        axis.title.x = element_text(size = 10, face = "bold"))
  
age_sex_symptom_factor_plot <- ggarrange(combined_symptoms_plot, combined_factors_plot + font("x.text", size = 10),
                    ncol = 2, nrow = 1, labels = c("A", "B"),
                    common.legend = TRUE,
                    legend = "bottom")
age_sex_symptom_factor_plot

#png
ggsave(
  "combined.final_plot_post_hoc_SES.png",
  plot = age_sex_symptom_factor_plot,
  device = "png",
  path = paste0(plots_path,"age_and_sex/"),
  width = 12,
  height = 6
)
```

