---
title: Age and sex-related variability in the presentation of anxiety and depression
  symptoms in the GLAD Study - formatting and descriptives. 
author: "Katherine N Thompson"
date: "10/02/2020"
output:
  word_document:
    toc: yes
    toc_depth: '2'
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: no
    number_sections: no
    highlight: monochrome
    theme: cerulean
code_folding: show
html_notebook:
  theme: cerulean
toc: yes
---

# Set up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      comment = NA,
                      prompt = FALSE,
                      cache = FALSE)
```

Delete everything in your global environment
```{r Clear global environment, include=FALSE}
remove(list = ls())
```

Packages
```{r Load packages, include=FALSE}
library(knitr)
library(car)
library(psych)
library(lubridate)
library(broom)
library(summarytools)
library(data.table)
library(eeptools)
library(magrittr)
library(reshape2)
library(lavaan)
library(mlogit)
library(foreign)
library(polycor)
library(REdaS)
library(scales)
library(forcats)
library(GPArotation)
library(tidyverse)
```

Retrieve the recent date
```{r Recent date, include=FALSE}
date = Sys.Date()
```

Define colours for plotting
```{r GLAD colour palette , include=FALSE}
palette = c(
  "#efc00b", #female
  "#b7dee8"  #male
  )

palette3 <- c("#78D9C5","#F5BE5E","#EEB6E9")
```

```{r source data path file}
source("../data_paths.R")
```

Read in demographics
```{r read in demographic data, include=FALSE}
data.dem.raw <- readRDS(paste0(data.raw_path, "DEM.rds"))
dim(data.dem.raw)
colnames(data.dem.raw)
data.dem.raw$birthyear <- NULL
# data.dem.raw$startdate <- NULL
data.dem.raw$enddate <- NULL
colnames(data.dem.raw)
```

```{r check start dates}
min(data.dem.raw$startdate)
max(data.dem.raw$startdate)
```


Read in self report mental health disorders
```{r Read in self report mental health disorders, include=FALSE}
data.mhd.raw <- readRDS(paste0(data.raw_path, "MHD.rds"))
dim(data.mhd.raw)
colnames(data.mhd.raw)
data.mhd.raw$sex <- NULL
data.mhd.raw$age <- NULL
data.mhd.raw$birthyear <- NULL
data.mhd.raw$startdate <- NULL
data.mhd.raw$enddate <- NULL
colnames(data.mhd.raw)
```

Read in PHQ9
```{r Read in PHQ, include=FALSE}
data.phq.raw <- readRDS(paste0(data.raw_path, "PHQ.rds"))
dim(data.phq.raw)
colnames(data.phq.raw)
data.phq.raw$sex <- NULL
data.phq.raw$age <- NULL
data.phq.raw$birthyear <- NULL
data.phq.raw$startdate <- NULL
data.phq.raw$enddate <- NULL
colnames(data.phq.raw)
```

Read in GAD7
```{r Read in GAD7, include=FALSE}
data.gad.raw <- readRDS(paste0(data.raw_path, "GAD.rds"))
dim(data.gad.raw)
colnames(data.gad.raw)
data.gad.raw$sex <- NULL
data.gad.raw$age <- NULL
data.gad.raw$birthyear <- NULL
data.gad.raw$startdate <- NULL
data.gad.raw$enddate <- NULL
colnames(data.gad.raw)
```

## Prepare item lists
```{r prepare item groups, include=FALSE}
# All the items in numeric form
ALL.items.full_numeric <- c("phq.anhedonia_numeric",
             "phq.depressed_mood_numeric",
             "phq.sleep_change_numeric",
             "phq.little_energy_numeric",
             "phq.eating_problems_numeric",
             "phq.worthlessness_numeric",
             "phq.concentration_problems_numeric",
             "phq.motor_problems_numeric",
             "phq.suicide_ideation_numeric",
             "gad.feeling_anxious_numeric",
             "gad.control_worrying_numeric",
             "gad.worry_toomuch_numeric",
             "gad.trouble_relaxing_numeric", 
             "gad.restless_numeric", 
             "gad.irritable_numeric",
             "gad.feeling_afraid_numeric")

# All the items as factors
ALL.items.full <- c("phq.anhedonia",
             "phq.depressed_mood",
             "phq.sleep_change",
             "phq.little_energy",
             "phq.eating_problems",
             "phq.worthlessness",
             "phq.concentration_problems",
             "phq.motor_problems",
             "phq.suicide_ideation",
             "gad.feeling_anxious",
             "gad.control_worrying",
             "gad.worry_toomuch",
             "gad.trouble_relaxing", 
             "gad.restless", 
             "gad.irritable",
             "gad.feeling_afraid")

# Create vector with items that are included in total score
GAD.items = c("gad.feeling_anxious_numeric",
              "gad.control_worrying_numeric",
              "gad.worry_toomuch_numeric",
              "gad.trouble_relaxing_numeric",
              "gad.restless_numeric",
              "gad.irritable_numeric",
              "gad.feeling_afraid_numeric")

# Create vector with items that are included in total score
PHQ.items = c("phq.anhedonia_numeric",
              "phq.depressed_mood_numeric",
              "phq.sleep_change_numeric",
              "phq.little_energy_numeric",
              "phq.eating_problems_numeric",
              "phq.worthlessness_numeric",
              "phq.concentration_problems_numeric",
              "phq.motor_problems_numeric",
              "phq.suicide_ideation_numeric")

# list of categorical items
cat.items <- c("phq.anhedonia.cat",
             "phq.depressed_mood.cat",       
             "phq.sleep_change.cat",
             "phq.little_energy.cat",
             "phq.eating_problems.cat",
             "phq.worthlessness.cat",
             "phq.concentration_problems.cat",
             "phq.motor_problems.cat",
             "phq.suicide_ideation.cat",
             "gad.feeling_anxious.cat",
             "gad.control_worrying.cat",
             "gad.worry_toomuch.cat",
             "gad.trouble_relaxing.cat", 
             "gad.restless.cat", 
             "gad.irritable.cat",
             "gad.feeling_afraid.cat")

# categorical items labeled for endorsement plot 
cat.items.labelled <- c("phq.anhedonia.cat.labelled",
             "phq.depressed_mood.cat.labelled",       
             "phq.sleep_change.cat.labelled",
             "phq.little_energy.cat.labelled",
             "phq.eating_problems.cat.labelled",
             "phq.worthlessness.cat.labelled",
             "phq.concentration_problems.cat.labelled",
             "phq.motor_problems.cat.labelled",
             "phq.suicide_ideation.cat.labelled",
             "gad.feeling_anxious.cat.labelled",
             "gad.control_worrying.cat.labelled",
             "gad.worry_toomuch.cat.labelled",
             "gad.trouble_relaxing.cat.labelled", 
             "gad.restless.cat.labelled", 
             "gad.irritable.cat.labelled",
             "gad.feeling_afraid.cat.labelled")

# create data frame that doesn't include "worrying too much" item
ALL.items.full.withoutworry <- c("phq.anhedonia",
                                 "phq.depressed_mood",
                                 "phq.sleep_change",
                                 "phq.little_energy",
                                 "phq.eating_problems",
                                 "phq.worthlessness",
                                 "phq.concentration_problems",
                                 "phq.motor_problems",
                                 "phq.suicide_ideation",
                                 "gad.feeling_anxious",
                                 "gad.control_worrying",
                                 "gad.trouble_relaxing", 
                                 "gad.restless", 
                                 "gad.irritable",
                                 "gad.feeling_afraid")

# create data frame that doesn't include "concentration problems" item
ALL.items.full.withoutconcentration <- c("phq.anhedonia",
                                 "phq.depressed_mood",
                                 "phq.sleep_change",
                                 "phq.little_energy",
                                 "phq.eating_problems",
                                 "phq.worthlessness",
                                 "phq.motor_problems",
                                 "phq.suicide_ideation",
                                 "gad.feeling_anxious",
                                 "gad.control_worrying",
                                 "gad.trouble_relaxing", 
                                 "gad.restless", 
                                 "gad.irritable",
                                 "gad.feeling_afraid")

# create data frame that doesn't include "concentration problems" and "irritability" item
ALL.items.full.withoutirritability <- c("phq.anhedonia",
                                 "phq.depressed_mood",
                                 "phq.sleep_change",
                                 "phq.little_energy",
                                 "phq.eating_problems",
                                 "phq.worthlessness",
                                 "phq.motor_problems",
                                 "phq.suicide_ideation",
                                 "gad.feeling_anxious",
                                 "gad.control_worrying",
                                 "gad.trouble_relaxing", 
                                 "gad.restless",
                                 "gad.feeling_afraid")

# continuous variables for descriptives
continuous.variables <- c("PHQ.total.score", 
                          "GAD.total.score", 
                          "age")
```

PHQ and GAD scoring keys
Create vector with scoring key: If the item is reversed coded use -1. There needs to be the correct number of values for each item in the questionnaire.

```{r PHQ and GAD scoring keys, include=FALSE}
# PHQ
PHQ.n.items = 9 # total number of items of the questionnaire
PHQ.items.key <- c(1,1,1,1,1,1,1,1,1) # scoring key

# GAD
GAD.n.items = 7 # total number of items of the questionnaire
GAD.items.key <- c(1,1,1,1,1,1,1) # scoring key
```

# Missingness {.tabset .tabset-fade}

## PHQ Missingness NAs per person - count and percentages
```{r PHQ Missingness}
data.phq.raw$na.per.person.phq <- rowSums(is.na(data.phq.raw[,colnames(data.phq.raw) %in% PHQ.items]))
freq(data.phq.raw$na.per.person.phq)

data.phq.raw$miss_pc_total.phq <- data.phq.raw$na.per.person.phq/PHQ.n.items
freq(data.phq.raw$miss_pc_total.phq)
```

## GAD Missingness NAs per person - count and percentages
```{r GAD Missingness}
data.gad.raw$na.per.person.gad <- rowSums(is.na(data.gad.raw[,colnames(data.gad.raw) %in% GAD.items]))
freq(data.gad.raw$na.per.person.gad)

data.gad.raw$miss_pc_total.gad <- data.gad.raw$na.per.person.gad/GAD.n.items
freq(data.gad.raw$miss_pc_total.gad)
```

# Merge PHQ, GAD, demographics and mental health disorders 

Create a list of the data frames to join together 
```{r list of dataframes to merge}
dataframe_list <- list(
  data.dem.raw,
  data.mhd.raw,
  data.gad.raw,
  data.phq.raw
)
```

Join data frames
```{r join all datasets}
data.joined <- plyr::join_all(
  dataframe_list,
  by = "ID" # Alternatively you can join by several columns
  )
#remove duplicate cols
data.joined <- data.joined[, !duplicated(colnames(data.joined))]
```

# **Data preparation** {.tabset .tabset-fade}

## Subset of individuals who signed up during the pandemic 

Dates for the definition of prepandemic
```{r Potential dates for prepandemic definition}
# Date of the first COVID-19 case in the United Kingdom (UK)
date_first_case_uk_january_31 <- as.POSIXct("2020-01-31")

# Date for assumed higher awareness of COVID / Lockdown in Italy 4. March 2020 / first festivals cancelled
date_higher_awareness_march_1 <- as.POSIXct("2020-03-01")

# Date the UK went into lockdown
date_uk_lockdown_march_23 <- as.POSIXct("2020-03-23")
```

```{r Prepandemic definitions january_31}
data.joined <- data.joined %>%
  mutate(
    prepandemic_january_31_numeric =
      if_else(
        startdate > date_first_case_uk_january_31, # The GLAD sign up is dated before the cut off (> longer ago)
        true = 0, # Participant has not signed up before the pandemic,
        false = 1, # Participant has signed up before the pandemic,
        missing = NA_real_
      )
  )

data.joined <- data.joined %>%
  mutate(
    prepandemic_january_31 =
      recode_factor(
        prepandemic_january_31_numeric,
        "0" = "Sign up after 31st January",
        "1" = "Sign up before 31st January"
      )
  )

data.joined %>%
  freq(prepandemic_january_31)
```

```{r Prepandemic definitions march_1}
data.joined <- data.joined %>%
  mutate(
    prepandemic_march_1_numeric =
      if_else(
        startdate > date_higher_awareness_march_1, # The GLAD sign up is dated before the cut off (> longer ago)
        true = 0, # Participant has not signed up before the pandemic,
        false = 1, # Participant has signed up before the pandemic,
        missing = NA_real_
      )
  )

data.joined <- data.joined %>%
  mutate(
    prepandemic_march_1 =
      recode_factor(
        prepandemic_march_1_numeric,
        "0" = "Sign up after 1st March",
        "1" = "Sign up before 1st March"
      )
  )

dat %>%
  freq(prepandemic_march_1)
```

```{r Prepandemic definitions march_23}
data.joined <- data.joined %>%
  mutate(
    prepandemic_march_23_numeric =
      if_else(
        startdate > date_uk_lockdown_march_23, # The GLAD sign up is dated before the cut off (> longer ago)
        true = 0, # Participant has not signed up before the pandemic,
        false = 1, # Participant has signed up before the pandemic,
        missing = NA_real_
      )
  )

data.joined <- data.joined %>%
  mutate(
    prepandemic_march_23 =
      recode_factor(
        prepandemic_march_23_numeric,
        "0" = "Sign up after 23rd March",
        "1" = "Sign up before 23rd March"
      )
  )

data.joined %>%
  freq(prepandemic_march_23)
```

## Missing data

### Convert all -77, -88 and -99 / "Seen but not answered", "Don't know", and "Prefer not to answer" to NA
```{r convert all -77,-88 and -99 to NA}
#convert all numeric variables to NA
dat <- data.joined %>%
  dplyr::mutate_all(., ~na_if(., -88)) %>% 
  dplyr::mutate_all(., ~na_if(., -99)) %>%
  dplyr::mutate_all(., ~na_if(., -77))
```

### If data import is older [i.e. does not include enddate and startdate]
Change -88 (Don't know), -99 (Prefer not to answer), and -77 (Seen but not answered) to NA
```{r change missing data to NA}
dat <- data.joined %>%
  select(-c("startdate","enddate")) %>% #Dropping startdate and enddate in new rds files
  mutate_all(., ~na_if(., -88)) %>% 
  mutate_all(., ~na_if(., -99)) %>%
  mutate_all(., ~na_if(., -77)) %>%
  mutate_all(., ~na_if(., "Seen but not answered")) %>%
  mutate_all(., ~na_if(., "Don't know")) %>%
  mutate_all(., ~na_if(., "Prefer not to say"))
```

### Getting rid of empty factor levels
```{r Empty factor levels}
dat <- dat %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer"))
```

## Identify duplicates
```{r Identify duplicates}
dat$ID.dup <- duplicated(dat$ID)

summary(as.factor(dat$ID.dup))

data.dup <- dat %>%
  filter(ID.dup == TRUE)

dat <- dat %>%
  filter(ID.dup == FALSE)
```

## Remove age outlier
```{r age outlier check}
# Outlier limits
age_upper_limit = 100
age_lower_limit = 16

#apply age limits
age.unc_upper <- which(dat$age > age_upper_limit)
age.unc_upper
age.unc_lower <- which(dat$age < age_lower_limit)
age.unc_lower

# Number of participants with outlier values
age.unc_outliers <- append(age.unc_upper, age.unc_lower)
length(age.unc_outliers)

# Remove outliers
dat$age <- with(dat, ifelse(age > age_upper_limit | age < age_lower_limit, NA, age))
```

## Remove those without sex, age and incomplete PHQ/GAD items
```{r remove anyone who doesnt report sex or age}
dat <- dat[complete.cases(dat[,"sex"]),]
summary(dat$ID)
dat <- dat[complete.cases(dat[,"age"]),]
summary(dat$ID)
#have removed incomplete PHQ/GAD items for factor analyses and conducting factor scores
dat <- dat[complete.cases(dat[, ALL.items.full_numeric]),]
summary(dat$ID)
```
 - All participants reported their biological sex and one person had their age missing, as I had previously removed this (age of 290). 
 - 1,781 were dropped because they had missing PHQ or GAD items. 

## Categorise age into groups 
```{r define age groups}
#create categorical age groups per 10 years 
dat$age.group_numeric <- case_when(dat$age >= 16 & dat$age <= 25 ~ "1",
                                  dat$age >= 26 & dat$age <= 35 ~ "2",
                                  dat$age >= 36 & dat$age <= 45 ~ "3",
                                  dat$age >= 46 & dat$age <= 55 ~ "4",
                                  dat$age >= 56 & dat$age <= 65 ~ "5",
                                  dat$age >= 66 & dat$age <= 75 ~ "6",
                                  dat$age >= 76 & dat$age <= 85 ~ "7",
                                  dat$age >= 86 & dat$age <= 95 ~ "8")

#create age groups as factors 
dat$age.group.factor <- recode_factor(dat$age.group_numeric,
                                         "1" = "16 to 25 years",
                                         "2" = "26 to 35 years",
                                         "3" = "36 to 45 years",
                                         "4" = "46 to 55 years",
                                         "5" = "56 to 65 years",
                                         "6" = "66 to 75 years",
                                         "7" = "76 to 85 years",
                                         "8" = "86 to 95 years")
#have a look at the data
freq(dat$age.group.factor, cumul = F) 
```

## Create self-report diagnostic groups

```{r create self-report MDD or any anxiety group}
#Reported MDD or any anxiety diagnosis
dat$MDD_anxiety <- ifelse(dat$mhd.MDD_numeric == "1" | 
                            dat$mhd.GAD_numeric == "1" |
                            dat$mhd.social_anxiety_numeric == "1" | 
                            dat$mhd.specific_phobia_numeric == "1" | 
                            dat$mhd.agoraphobia_numeric == "1" | 
                            dat$mhd.panic_disorder_numeric == "1", 1, 0)
dat <- dat %>%
  mutate(
    MDD_anxiety.factor =
    recode_factor(MDD_anxiety, 
                  "1" = "MDD and any anxiety",
                  "0" = "No MDD and any anxiety"
    ))
table(dat$MDD_anxiety.factor)
```

```{r create self-report MDD only group}
#MDD only group - not reported any anxiety disorder
dat$MDD.only <- ifelse(dat$mhd.MDD_numeric == "1" & 
                            dat$mhd.GAD_numeric == "0" &
                            dat$mhd.social_anxiety_numeric == "0" & 
                            dat$mhd.specific_phobia_numeric == "0" & 
                            dat$mhd.agoraphobia_numeric == "0" & 
                            dat$mhd.panic_disorder_numeric == "0", 1, 0)
dat <- dat %>%
  mutate(
    MDD.only.factor =
    recode_factor(MDD.only, 
                  "1" = "MDD only",
                  "0" = "No MDD only"
    ))
table(dat$MDD.only.factor)
```

```{r create self-report GAD only group}
#GAD only group
dat$GAD.only <- ifelse(dat$mhd.GAD_numeric == "1" & 
                            dat$mhd.MDD_numeric == "0" &
                            dat$mhd.social_anxiety_numeric == "0" & 
                            dat$mhd.specific_phobia_numeric == "0" & 
                            dat$mhd.agoraphobia_numeric == "0" & 
                            dat$mhd.panic_disorder_numeric == "0", 1, 0)
dat <- dat %>%
  mutate(
    GAD.only.factor =
    recode_factor(GAD.only, 
                  "1" = "GAD only",
                  "0" = "No GAD only"
    ))
table(dat$GAD.only.factor)
```

```{r create self-report comorbid MDD and GAD only groups}
#Comorbid MDD and GAD
dat$MDD_GAD <- ifelse(dat$mhd.GAD_numeric == "1" & 
                            dat$mhd.MDD_numeric == "1" &
                            dat$mhd.social_anxiety_numeric == "0" & 
                            dat$mhd.specific_phobia_numeric == "0" & 
                            dat$mhd.agoraphobia_numeric == "0" & 
                            dat$mhd.panic_disorder_numeric == "0", 1, 0)
dat <- dat %>%
  mutate(
    MDD_GAD.factor =
    recode_factor(MDD_GAD, 
                  "1" = "Comorbid MDD and GAD only",
                  "0" = "No comorbid MDD and GAD only"
    ))
table(dat$MDD_GAD.factor)
```

```{r create self-report any other anxiety disorder group}
#All who have any other anxiety disorder
dat$anxiety.other.diagnosis <- ifelse(dat$mhd.social_anxiety_numeric == "1" | 
                                            dat$mhd.specific_phobia_numeric == "1" | 
                                            dat$mhd.agoraphobia_numeric == "1" | 
                                            dat$mhd.panic_disorder_numeric == "1", 1, 0)
dat <- dat %>%
  mutate(
    anxiety.other.diagnosis.factor =
    recode_factor(anxiety.other.diagnosis, 
                  "1" = "Other anxiety disorder",
                  "0" = "No other anxiety disorder"
    ))
table(dat$anxiety.other.diagnosis.factor)
```

```{r create self-report any other anxiety disorder only group}
#all who have any other anxiety disorder only 
dat$anxiety.other.diagnosis.only <- ifelse((dat$mhd.social_anxiety_numeric == "1" | 
                                            dat$mhd.specific_phobia_numeric == "1" | 
                                            dat$mhd.agoraphobia_numeric == "1" | 
                                            dat$mhd.panic_disorder_numeric == "1" &
                                            (dat$mhd.GAD_numeric == "0") &  
                                            dat$mhd.MDD_numeric == "0"), 1, 0)
dat <- dat %>%
  mutate(
    anxiety.other.diagnosis.only.factor =
    recode_factor(anxiety.other.diagnosis.only, 
                  "1" = "Other anxiety disorder without MDD and GAD",
                  "0" = "No other anxiety disorder without MDD and GAD"
    ))
table(dat$anxiety.other.diagnosis.only.factor)
```

## Create data sets with just men/women
```{r datasets for males and males}
data.female <- dat %>% filter(sex == "Female")
data.male <- dat %>% filter (sex == "Male")
```

## Create Highest Education variable

Assumes Degree > A level > NVQ > CSE/GCSE
```{r define highest education}
#create numeric version of the highest education variable
dat <- dat %>%
  mutate(
    highest_education_numeric =
      case_when(
        dem.university == "Yes" ~ 4,
        dem.alevels == "Yes" ~ 3,
        dem.NVQ == "Yes" ~ 2,
        dem.gcse == "Yes" ~ 1,
        dem.cse == "Yes" ~ 1))

#recode the numeric version into a factor
dat <- dat %>%
  mutate(
    highest_education =
      recode_factor(highest_education_numeric,
        `1` = "GCSE/CSE",
        `2` = "NVQ",
        `3` = "A-levels",
        `4` = "University"))

freq(dat$highest_education)
```

## Categroise PHQ-9 and GAD-7 items

### PHQ
```{r making PHQ items categorical}
#categorise the PHQ items for the logistic regressions
dat$phq.depressed_mood.cat <- ifelse(dat$phq.depressed_mood_numeric > 0, 1, 0)
dat$phq.anhedonia.cat <- ifelse(dat$phq.anhedonia_numeric > 0, 1, 0)
dat$phq.sleep_change.cat <- ifelse(dat$phq.sleep_change_numeric > 0, 1, 0)
dat$phq.little_energy.cat <- ifelse(dat$phq.little_energy_numeric > 0, 1, 0)
dat$phq.eating_problems.cat <- ifelse(dat$phq.eating_problems_numeric > 0, 1, 0)
dat$phq.worthlessness.cat <- ifelse(dat$phq.worthlessness_numeric > 0, 1, 0)
dat$phq.concentration_problems.cat <- ifelse(dat$phq.concentration_problems_numeric > 0, 1, 0)
dat$phq.motor_problems.cat <- ifelse(dat$phq.motor_problems_numeric > 0, 1, 0)
dat$phq.suicide_ideation.cat <- ifelse(dat$phq.suicide_ideation_numeric > 0, 1, 0)

table(dat$phq.depressed_mood.cat)
```

### GAD
```{r making GAD items categorical}
#categorise the GAD items for the logistic regressions
dat$gad.feeling_anxious.cat <- ifelse(dat$gad.feeling_anxious_numeric > 0, 1, 0)
dat$gad.control_worrying.cat <- ifelse(dat$gad.control_worrying_numeric > 0, 1, 0)
dat$gad.worry_toomuch.cat <- ifelse(dat$gad.worry_toomuch_numeric > 0, 1, 0)
dat$gad.trouble_relaxing.cat <- ifelse(dat$gad.trouble_relaxing_numeric > 0, 1, 0)
dat$gad.restless.cat <- ifelse(dat$gad.restless_numeric > 0, 1, 0)
dat$gad.irritable.cat <- ifelse(dat$gad.irritable_numeric > 0, 1, 0)
dat$gad.feeling_afraid.cat <- ifelse(dat$gad.feeling_afraid_numeric > 0, 1, 0)

table(dat$gad.feeling_anxious.cat)
```

## Total scores

### PHQ
```{r calculate PHQ total score }
#calulate total score
PHQ.scored.items <- scoreItems(keys = PHQ.items.key,
                                        items = dat[PHQ.items],
                                        totals = TRUE, 
                                        missing = TRUE, 
                                        impute = 'none', #this should not matter as the data is full
                                        min = 0, 
                                        max = 3)
#Add in column to data frame
dat$PHQ.total.score <- PHQ.scored.items$scores

table(dat$PHQ.total.score)
```

### GAD
```{r calculate GAD total score }
#calulate total score
GAD.scored.items <- scoreItems(keys = GAD.items.key,
                                        items = dat[GAD.items],
                                        totals = TRUE, 
                                        missing = TRUE, 
                                        impute = 'none', 
                                        min = 0, 
                                        max = 3)
#Add in column to data frame
dat$GAD.total.score <- GAD.scored.items$scores

table(dat$GAD.total.score)
```

## Create ordered factors for PHQ anf GAD items to aid labelling for endorsement plot

```{r create ordered factors for endorsement plot}
dat$phq.anhedonia.cat.labelled <- recode_factor(dat$phq.anhedonia.cat, 
                                          "0" = "Did not report symptom", 
                                           "1" = "Reported symptom")

dat$phq.depressed_mood.cat.labelled <- recode_factor(dat$phq.depressed_mood.cat, 
                                           "0" = "Did not report symptom", 
                                           "1" = "Reported symptom")

dat$phq.suicide_ideation.cat.labelled <- recode_factor(dat$phq.suicide_ideation.cat, 
                                          "0" = "Did not report symptom", 
                                           "1" = "Reported symptom")

dat$phq.concentration_problems.cat.labelled <- recode_factor(dat$phq.concentration_problems.cat, 
                                           "0" = "Did not report symptom", 
                                           "1" = "Reported symptom")

dat$phq.worthlessness.cat.labelled <- recode_factor(dat$phq.worthlessness.cat, 
                                           "0" = "Did not report symptom", 
                                           "1" = "Reported symptom")

dat$phq.little_energy.cat.labelled <- recode_factor(dat$phq.little_energy.cat, 
                                           "0" = "Did not report symptom", 
                                           "1" = "Reported symptom")

dat$phq.eating_problems.cat.labelled <- recode_factor(dat$phq.eating_problems.cat, 
                                           "0" = "Did not report symptom", 
                                           "1" = "Reported symptom")

dat$phq.sleep_change.cat.labelled <- recode_factor(dat$phq.sleep_change.cat, 
                                           "0" = "Did not report symptom", 
                                           "1" = "Reported symptom")

dat$phq.motor_problems.cat.labelled <- recode_factor(dat$phq.motor_problems.cat, 
                                          "0" = "Did not report symptom", 
                                           "1" = "Reported symptom")

dat$gad.worry_toomuch.cat.labelled <- recode_factor(dat$gad.worry_toomuch.cat, 
                                           "0" = "Did not report symptom", 
                                           "1" = "Reported symptom")

dat$gad.control_worrying.cat.labelled <- recode_factor(dat$gad.control_worrying.cat, 
                                           "0" = "Did not report symptom", 
                                           "1" = "Reported symptom")

dat$gad.feeling_anxious.cat.labelled <- recode_factor(dat$gad.feeling_anxious.cat, 
                                           "0" = "Did not report symptom", 
                                           "1" = "Reported symptom")

dat$gad.feeling_afraid.cat.labelled <- recode_factor(dat$gad.feeling_afraid.cat, 
                                           "0" = "Did not report symptom", 
                                           "1" = "Reported symptom")

dat$gad.trouble_relaxing.cat.labelled <- recode_factor(dat$gad.trouble_relaxing.cat, 
                                           "0" = "Did not report symptom", 
                                           "1" = "Reported symptom")

dat$gad.restless.cat.labelled <- recode_factor(dat$gad.restless.cat, 
                                           "0" = "Did not report symptom", 
                                           "1" = "Reported symptom")

dat$gad.irritable.cat.labelled <- recode_factor(dat$gad.irritable.cat, 
                                           "0" = "Did not report symptom", 
                                           "1" = "Reported symptom")
```

# **Sample descriptives**

## Demographics
```{r call packages for descriptives, include=FALSE}
library(janitor)
library(tidyr)
library(tidyverse)
```

```{r descriptives for age}
continuous_age_descriptives_raw <- dat %>%
  group_by(sex) %>%
  summarise(mean = mean(age), sd = sd(age), min = min(age), max = max(age)) %>%
  pivot_longer(cols = c("mean", "sd", "min", "max")) %>%
  rename(descriptive = name) %>%
  pivot_wider(names_from = sex, values_from = value) %>%
  mutate(Total = 
           c(mean(dat$age), sd(dat$age), min(dat$age), max(dat$age)),
         sd_male = c(sd(data.male$age), NA_real_, NA_real_, NA_real_),
         sd_female = c(sd(data.female$age), NA_real_, NA_real_, NA_real_),
         sd_total = c(sd(dat$age), NA_real_, NA_real_, NA_real_)) %>%
  select(
    "Descriptive" = descriptive,
    "Male" = Male,
    "% Male" = sd_male,
    "Female" = Female,
    "% Female" = sd_female,
    "Total" = Total,
    "% Total" = sd_total) 
  
continuous_age_descriptives <- continuous_age_descriptives_raw[-c(2), ]
```

```{r descriptives for age group}
age_group_descriptives <- dat %>% #create object to bind later on
  count(sex, age.group.factor) %>%
  mutate(Prop = round(n/sum(n)*100, 2)) %>%
  pivot_wider(names_from = sex, values_from = c(n, Prop)) %>%
  mutate(Total = n_Male + n_Female,
         Prop_total = Prop_Male + Prop_Female) %>% 
  select(
    "Descriptive" = age.group.factor,
    "Male" = n_Male,
    "% Male" = Prop_Male,
    "Female" = n_Female,
    "% Female" = Prop_Female,
    "Total" = Total,
    "% Total" = Prop_total)
age_group_descriptives 
```

```{r descriptives for highest education}
highest_education_descriptives.raw <- dat %>% #create object to bind later on
  count(sex, highest_education) %>%
  mutate(Prop = round(n/sum(n)*100, 2)) %>%
  pivot_wider(names_from = sex, values_from = c(n, Prop)) %>%
  mutate(Total = n_Male + n_Female,
         Prop_total = Prop_Male + Prop_Female) %>% 
  select(
    "Descriptive" = highest_education,
    "Male" = n_Male,
    "% Male" = Prop_Male,
    "Female" = n_Female,
    "% Female" = Prop_Female,
    "Total" = Total,
    "% Total" = Prop_total)
highest_education_descriptives <- highest_education_descriptives.raw[-c(5),]
```

```{r descriptives for self-report diagnosis}
#MDD total
diagnosis_descriptives_mdd.raw <- dat %>% #create object to bind later on
  count(sex, mhd.MDD) %>%
  mutate(Prop = round(n/sum(n)*100, 2)) %>%
  pivot_wider(names_from = sex, values_from = c(n, Prop)) %>%
  mutate(Total = n_Male + n_Female,
         Prop_total = Prop_Male + Prop_Female) %>% 
  select(
    "Descriptive" = mhd.MDD,
    "Male" = n_Male,
    "% Male" = Prop_Male,
    "Female" = n_Female,
    "% Female" = Prop_Female,
    "Total" = Total,
    "% Total" = Prop_total)

diagnosis_descriptives_mdd <- diagnosis_descriptives_mdd.raw[-c(1,3), ] 

#mdd only 
diagnosis_descriptives_mdd_only.raw <- dat %>% #create object to bind later on
  count(sex, MDD.only.factor) %>%
  mutate(Prop = round(n/sum(n)*100, 2)) %>%
  pivot_wider(names_from = sex, values_from = c(n, Prop)) %>%
  mutate(Total = n_Male + n_Female,
         Prop_total = Prop_Male + Prop_Female) %>% 
  select(
    "Descriptive" = MDD.only.factor,
    "Male" = n_Male,
    "% Male" = Prop_Male,
    "Female" = n_Female,
    "% Female" = Prop_Female,
    "Total" = Total,
    "% Total" = Prop_total)

diagnosis_descriptives_mdd_only <- diagnosis_descriptives_mdd_only.raw[-c(2,3), ]

#GAD
diagnosis_descriptives_gad.raw <- dat %>% #create object to bind later on
  count(sex, mhd.GAD) %>%
  mutate(Prop = round(n/sum(n)*100, 2)) %>%
  pivot_wider(names_from = sex, values_from = c(n, Prop)) %>%
  mutate(Total = n_Male + n_Female,
         Prop_total = Prop_Male + Prop_Female) %>% 
  select(
    "Descriptive" = mhd.GAD,
    "Male" = n_Male,
    "% Male" = Prop_Male,
    "Female" = n_Female,
    "% Female" = Prop_Female,
    "Total" = Total,
    "% Total" = Prop_total)

diagnosis_descriptives_gad <- diagnosis_descriptives_gad.raw[-c(1,3), ] 

#GAD only
diagnosis_descriptives_gad_only.raw <- dat %>% #create object to bind later on
  count(sex, GAD.only.factor) %>%
  mutate(Prop = round(n/sum(n)*100, 2)) %>%
  pivot_wider(names_from = sex, values_from = c(n, Prop)) %>%
  mutate(Total = n_Male + n_Female,
         Prop_total = Prop_Male + Prop_Female) %>% 
  select(
    "Descriptive" = GAD.only.factor,
    "Male" = n_Male,
    "% Male" = Prop_Male,
    "Female" = n_Female,
    "% Female" = Prop_Female,
    "Total" = Total,
    "% Total" = Prop_total)

diagnosis_descriptives_gad_only <- diagnosis_descriptives_gad_only.raw[-c(2,3), ]

#Other anxiety diagnoses
diagnosis_descriptives_other_anxiety.raw <- dat %>% #create object to bind later on
  count(sex, anxiety.other.diagnosis.factor) %>%
  mutate(Prop = round(n/sum(n)*100, 2)) %>%
  pivot_wider(names_from = sex, values_from = c(n, Prop)) %>%
  mutate(Total = n_Male + n_Female,
         Prop_total = Prop_Male + Prop_Female) %>% 
  select(
    "Descriptive" = anxiety.other.diagnosis.factor,
    "Male" = n_Male,
    "% Male" = Prop_Male,
    "Female" = n_Female,
    "% Female" = Prop_Female,
    "Total" = Total,
    "% Total" = Prop_total)

diagnosis_descriptives_other_anxiety <- diagnosis_descriptives_other_anxiety.raw[-c(2,3), ]

#Other anxiety disorder without MDD and GAD
diagnosis_descriptives_other_anxiety.only.raw <- dat %>% #create object to bind later on
  count(sex, anxiety.other.diagnosis.only.factor) %>%
  mutate(Prop = round(n/sum(n)*100, 2)) %>%
  pivot_wider(names_from = sex, values_from = c(n, Prop)) %>%
  mutate(Total = n_Male + n_Female,
         Prop_total = Prop_Male + Prop_Female) %>% 
  select(
    "Descriptive" = anxiety.other.diagnosis.only.factor,
    "Male" = n_Male,
    "% Male" = Prop_Male,
    "Female" = n_Female,
    "% Female" = Prop_Female,
    "Total" = Total,
    "% Total" = Prop_total)

diagnosis_descriptives_other_anxiety.only <- diagnosis_descriptives_other_anxiety.only.raw[-c(2,3), ]

#bind all rows together
diagnosis_descriptives <- rbind(diagnosis_descriptives_mdd, 
                              diagnosis_descriptives_mdd_only, 
                              diagnosis_descriptives_gad, 
                              diagnosis_descriptives_gad_only, 
                              diagnosis_descriptives_other_anxiety,
                              diagnosis_descriptives_other_anxiety.only)
```

```{r descriptives for ethnicity }
ethnicity_descriptives.raw <- dat %>% #create object to bind later on
  count(sex, dem.ethnicity) %>%
  mutate(Prop = round(n/sum(n)*100, 2)) %>%
  pivot_wider(names_from = sex, values_from = c(n, Prop)) %>%
  mutate(Total = n_Male + n_Female,
         Prop_total = Prop_Male + Prop_Female) %>% 
  select(
    "Descriptive" = dem.ethnicity,
    "Male" = n_Male,
    "% Male" = Prop_Male,
    "Female" = n_Female,
    "% Female" = Prop_Female,
    "Total" = Total,
    "% Total" = Prop_total)

ethnicity_descriptives <- ethnicity_descriptives.raw[-c(7), ]
ethnicity_descriptives
```

```{r bind all descriptives together}
descriptives <- bind_rows(
  continuous_age_descriptives,
  age_group_descriptives,
  ethnicity_descriptives,
  highest_education_descriptives,
  diagnosis_descriptives
)

#add in column to show which descriptives they are
descriptives <- descriptives %>%
  mutate(
    Category = 
      c("Age", "Age", "Age",
        "Age group", "Age group", "Age group", "Age group", "Age group", "Age group", "Age group", "Age group",
        "Ethnicity", "Ethnicity", "Ethnicity", "Ethnicity", "Ethnicity", "Ethnicity",
        "Highest education level", "Highest education level", "Highest education level", "Highest education level",
        "Self-reported diagnosis", "Self-reported diagnosis", "Self-reported diagnosis", "Self-reported diagnosis", "Self-reported diagnosis", "Self-reported diagnosis"),
    Male = round(Male, 0),
    Female = round(Female, 0),
    Total = round(Total, 0),
    `% Male /SD`= round(`% Male`, 2),
    `% Female /SD`= round(`% Female`, 2),
    `% Total`= round(`% Total`, 2)) %>%
select(
    Category,
    Descriptive,
    Male,
    `% Male /SD`,
    Female,
    `% Female /SD`,
    Total,
    `% Total`) 

#final table to be viewed in word
kable(descriptives) 
```

```{r RE create datasets for males and males, include=FALSE}
data.female <- dat %>% filter(sex == "Female")
data.male <- dat %>% filter (sex == "Male")
```

## Continuous descriptives
```{r continuous descriptives for age}
library(tidyverse)
dat %>%
  group_by(sex) %>%
  descr(dat[,continuous.variables],
      round.digits = 2,
      transpose = FALSE)

descr(x = dat[,continuous.variables],
      round.digits = 2,
      transpose = TRUE)
```

## Individuals that meet current cut off criteria for GAD-7 and PHQ-9

* PHQ criteria: Those who score above 9
* GAD criteria: Those who score above 7

```{r how many meet current cut off}
#how many meet the cinical cut-off for PHQ
meetPHQcutoff <- ifelse(dat$PHQ.total.score > 9, 1, 0)
table(meetPHQcutoff)

#how many meet the cinical cut-off for GAD
meetGADcutoff <- ifelse(dat$GAD.total.score > 6, 1, 0)
table(meetGADcutoff)
```

## Endorsement of items: likert scale plot 
```{r likert plot of PHQ and GAD items}
endorsement_plot <- sjPlot::plot_likert(
  items = dat[,cat.items.labelled],
  title = "Endorsement of PHQ-9 and GAD-7 symptoms in GLAD",
  axis.labels = c("Anhedonia",
             "Depressed mood",
             "Sleep change",
             "Little energy",
             "Weight or Appetite problems",
             "Worthlessness",
             "Concentration problems",
             "Motor problems",
             "Suicide ideation",
             "Feeling anxious",
             "Difficulty controlling worrying",
             "Worry too much",
             "Trouble relaxing", 
             "Restlessness", 
             "Irritable",
             "Feeling afraid"),
  wrap.labels = 20,
  digits = 0,
  reverse.scale = TRUE,
  cat.neutral = NULL,
  value = "show",
  catcount = 2,
  geom.colors = c("#b7dee8","#efc00b"))

#have a look at the plot
endorsement_plot

#png
ggsave(
  "endorsement_plot.png",
  plot = endorsement_plot,
  device = "png",
  path = paste0(plots_path, "endorsement_plot/"),
  width = 8,
  height = 9
)
```

# Save data

```{r save data file}
# save R data file
saveRDS(object = dat, file = paste0(data_path, "data_formatted.rds"))
```





